---
output: html_document
editor_options: 
  chunk_output_type: console
---
## Getting Started

First load the __ggplot2__ package
```{r, , message = FALSE, echo = TRUE}
library("ggplot2")
```
and the OK Cupid data set
```{r, , echo = TRUE}
data(okcupid, package = "jrSouth")
```

## Scatter plots

1. Let's start by creating a basic scatter plot of the heights and ages. We do this using the `geom_point()` command

```{r, , fig.keep='none', echo = TRUE}
# alpha makes the points transparent
ggplot(data = okcupid) + 
  geom_point(aes(x = age, y = height), alpha = 0.2)
```

2. To save typing, we're going to store the original ggplot object as a variable

```{r, , fig.keep='none', echo = TRUE}
g = ggplot(data = okcupid)
g1 = g + geom_point(aes(x = age, y = height), alpha = 0.2)
```

So now running `g1` will produce the graph
```{r, , fig.keep='none', echo = TRUE}
g1
```

3. The arguments `x` and `y` here are called aesthetics. What do you think happens if you omit the `y` aesthetic?

```{r, }
# this gives an error
```

4. For `geom_point()`, both the `x` and `y` aesthetics are required. This particular geom has other aesthetics: shape, colour, size and alpha.^[These are available for most geoms. For a collection of aesthetics see the relevant help pages.]  For instance we can specify that we wish to map the variable `sex` to a colour aesthetic by including it inside `aes()`

    ```{r, , fig.keep='none', echo = TRUE}
    g + geom_point(aes(x = age, y = height, colour = sex))
    ```

5. Change `colour = sex` to `colour = height`. Why do you think there's a change in the legend?

## Bar plots

`geom_bar()` can be used to create a bar chart. It requires only one aesthetic and that is `x`. For the provided aesthetic, the frequencies will be calculated and shown as a bar. For example

```{r, , fig.keep='none', echo = TRUE}
g + geom_bar(aes(x = body_type))
```

1. Change the axis labels to "Body Type" and "Total" by adding two more layers, 
using `+ xlab("Body Type")` and `+ ylab("Total")`

```{r, , fig.keep='none'}
g + geom_bar(aes(x = body_type)) + 
  xlab("Body Type") + 
  ylab("Total")
```

2. Split the graph up into the two genders (hint: use `colour` and `fill`). 

```{r, , fig.keep='none'}
# What happens if you only have colour or only fill?
g + geom_bar(aes(x = body_type, colour = sex, fill = sex))
```

3. With such long labels, it might make more sense to rotate the graph such that the bars and labels are horizontal. Switch the `x` and `y` axes using a `coord_flip()` layer added to your graph.

```{r, , fig.keep='none'}
g + geom_bar(aes(x = body_type, colour = sex, fill = sex)) + 
  coord_flip()
```

4. I am not too keen on how the female and male bars are displayed on top of each other. The argument to change these is `position`.  The default is `stack`, for example we can put the bars next to each other using 

    ```{r, , fig.keep='none', echo = TRUE}
    g + geom_bar(aes(x = body_type, colour = sex, fill = sex), position = "dodge") + 
      coord_flip()
    ```

    Other values you might try here are `position = fill`, `position = jitter`, or `position = identity`. 

5. What does the `fill` position argument do?

```{r, , fig.keep = "none"}
g + geom_bar(aes(x = body_type, colour = sex, fill = sex), position = "fill") + 
  coord_flip()
# puts the values on a common scale (all sum to 1)
```

## Box plots

1. Box plots are a great way to visualise the shape of a distribution of some variable. Start by creating a boxplot of peoples ages in the `okcupid` data. The `x = 1` in the code below lets us have just a single boxplot for all ages 

```{r, , echo = TRUE, out.width="70%"}
g = ggplot(okcupid) 
g + geom_boxplot(aes(x = 1, y = age))
```

2. Switch out the `x = 1` for `x = smokes` to get a boxplot for each group

```{r, }
g + geom_boxplot(aes(x = smokes, y = age))
```

## Copy cat


```{r, echo=FALSE, message=FALSE, fig.keep='none'}
data(mpg, package="ggplot2")
# mpg$drv = as.character(mpg$drv)
# mpg[mpg$drv == "f",]$drv = "Front"
# mpg[mpg$drv == "r",]$drv = "Rear"
# mpg[mpg$drv == "4",]$drv = "4wd"
# mpg$drv = factor(mpg$drv,
#    levels = c("Front", "Rear", "4wd"))

g = ggplot(data=mpg, aes(x=displ, y=hwy))
g1 = g + geom_point() + stat_smooth(linetype=2) +
  xlab("Displacement") + ylab("Highway mpg")
g2 = g + geom_point() + stat_smooth(aes(colour=drv))
```

```{r, out.width='0.5\\textwidth', message=FALSE, label = "F1", fig.cap = "Graphics for section 1", echo = FALSE, fig.keep='all'}
g1
```


The aim of this section is to recreate the graphics in figure 1 and 2. Feel free to experiment. To begin, load the package

```{r, message=FALSE, echo = TRUE}
library("ggplot2")
```

\noindent and the `mpg` data set

```{r, results='hide', echo = TRUE}
data(mpg, package="ggplot2")
dim(mpg)
```

1) Figure 1: Create a scatter plot of engine displacement,
  `displ`, against highway mpg, `hwy`. To get started:

```{r, fig.keep='none', tidy=FALSE, echo = TRUE}
ggplot(data=mpg, aes(x=displ, y=hwy)) +
  geom_point() + xlab("Displacement")
```

Now add a dashed loess line and change the $y$-axis label.
Hint: try `stat_smooth()` and `ylab('New label')`.


```{r, label = "F2", out.width='0.5\\textwidth', echo=FALSE, message=FALSE, fig.cap="Graphics for section 1", fig.keep='all'}
g2
```
```{r, message=FALSE, fig.keep='none'}
g1 = g + geom_point() + stat_smooth(linetype=2) +
  xlab("Displacement") + ylab("Highway mpg")
```
2) Figure 2: Using `stat_smooth()`, add a loess line conditional
  on the drive.
  
```{r, message=FALSE, eval = FALSE}
g2 = g + geom_point() + stat_smooth(aes(colour=drv))
```

